---
title: "Data visualisation in R"
#format:
 # html:
 #   grid: 
  #    body-width: 2000px
   #   sidebar-width: 200px
    #  margin-width: 200px
code-fold: show
code-link: true
code-tools: true
highlight-style: github
bibliography: references.bib
author: 
  - name: Dr Nipun Shrestha
    url: https://ctc.usyd.edu.au/about-us/our-people/academics-research-fellows/dr-nipun-shrestha/
    orcid: 0000-0003-3542-8130
    degrees: 
      - MBBS
      - MPH
      - PhD
google-scholar: true
lightbox: true
engine: knitr
comments:
    utterances: 
      repo: Nipun-Shrestha/Profile_page

#filters:
 # - social-share
#share:
 # permalink: https://nipun-shrestha.github.io/Profile_page/Tutorial.html
  #description: "awesome page"
  #twitter: true
  #facebook: true
  #linkedin: true
---

# Age-standardized DALYs rate of cardiovascular disease per 100,000 population from 2001 to 2021 in South Asia by country’s GDP/capita (US\$)

The GDP per capita (inflation-adjusted) are plotted in x-axis and corresponding values of age-standardized cardiovascular disease (CVD) rates for A) Females and B) Males are plotted in y-axis. The size of bubble corresponds to the size of the population of the country. The definition of cardiovascular disease includes all eleven CVD types and is described [elsewhere](https://www.healthdata.org/research-analysis/gbd "GBD"). The data on cardiovascular disease rates were obtained from the [Global Burden of Disease](https://www.healthdata.org/research-analysis/gbd "GBD") and GDP per capita from [World Bank.](https://data.worldbank.org/indicator/NY.GDP.PCAP.CD?end=2022&locations=8S&start=2001)

::: panel-tabset
#### Loading Packages

```{r liby, warning = FALSE, message = FALSE}
library(conflicted)
library(tidyverse)
library(ggplot2)
library(gganimate)
theme_set(theme_bw())
library("gifski")
```

#### Loading Data

```{r dat, warning = FALSE, message = FALSE}
load("~/GitHub/Profile/GBD.Rdata")
dat1 <- dat1 %>% dplyr::filter(measure =="DALYs (Disability-Adjusted Life Years)") 
```
:::

#### Plotting graph

```{r dpi=400}

p <- ggplot(
dat1,
aes(x = income, y=val, size = Population, colour = Country)
) +
geom_point(show.legend = TRUE, alpha = 1) +
scale_color_viridis_d() +
scale_size(range = c(2, 12)) +
scale_x_log10() +
labs(x = "Income per person (GDP/capita PPPS inflation-adjusted)", y = "Age-standardized DALYs rate per 100,000") +
  facet_wrap(~sex)


R <- p + transition_time(year) +
labs(title = "Year: {frame_time}") +
shadow_mark(alpha = 2, size = 2) +
  facet_wrap(~sex)
R
#animate(R, height=400, width=800, renderer=gifski_renderer())

```

# Creating Maps and Mapping data with ggplot2

::: callout-note
The shape file for Nepal can be obtained from Government of [Government of Nepal geoportalal](https://nationalgeoportal.gov.np/#/map)and availability of human resource data from [Open data Nepal](https://opendatanepal.com/dataset?category=Health&page=4).
:::

::: panel-tabset
#### Loading Packages

```{r, warning = FALSE, message = FALSE}
library (sf)
library (ggplot2)
library (viridis)
library (ggthemes)
library (tidyverse)
```

#### Loading & Preparing Data

```{r, warning = FALSE, message = FALSE}

#Loading datasets

data <- st_read("~/GitHub/Profile/districts.shp")
medical_doc <- read.csv("~/GitHub/Profile/medical_doc.csv")

#Preparing dataset

##Merging shape data with human resource data
df <- medical_doc %>% mutate(across(where(is.character), toupper))
dat <- merge(data, df, by="DISTRICT")

#Creating list of co-ordinates for adding district names in the map
points <- cbind(dat, st_coordinates(st_centroid(dat$geometry)))

```
:::

#### Plotting map and mapping data

```{r fig.height= 15, fig.width=17, warning = FALSE, message = FALSE}


# Plotting data

ggplot(data=dat) +geom_sf (aes(fill=Fulfilled_percent), color="black", size=0.2) + 
    scale_fill_viridis_c(option = "viridis", trans="sqrt")+ 
    geom_text(data=points, aes(x=X, y=Y, label= paste(DISTRICT)), color="darkblue", size = 2.5) + 
    ggtitle("Fullfilled percentage of sanctioned post of Medical doctors in Nepal 2072/73") +
    theme_void()

```
